---
title: "Reproducible Research: Peer Assessment 1"
author: "Nazmi Anik"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data  

1. Unzip the file first, if the unzipped file is not already in the working directory.  
```{r, echo = TRUE}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
```

2. Load the data ( i.e. read.csv() )  
```{r, echo = TRUE}
#read the table
activity <- read.csv("activity.csv", stringsAsFactors=FALSE, na.strings="NA")
```

3. Process/transform the data (if necessary) into a format suitable for your analysis  
```{r, echo = TRUE}
#change the date column's class from string to Date
activity[,"date"] <- as.Date(activity[,"date"],format='%Y-%m-%d')
#display the first few rows
head(activity)
```

## What is mean total number of steps taken per day?  

1. Calculate the total number of steps taken per day  
```{r, message=FALSE}
#import dplyr library
library(dplyr)
```
```{r, echo=TRUE}
#group the activity table by day
activityDay <- group_by(activity, date)
#summarize this grouped table with the sum of steps for each day
stepDay <- summarize(activityDay, steps_by_day = sum(steps))
#For this part of the assignment, you can ignore the missing values in the dataset.
stepDay <- na.omit(stepDay)
#display the first few rows
head(stepDay)
```

2. Make a histogram of the total number of steps taken each day  
```{r, echo=TRUE}
hist(stepDay$steps_by_day, main="Number of Steps Taken Each Day", xlab="Number of Steps per Day", ylab="Frequency", col = "red")
```

3. Calculate and report the mean and median of the total number of steps taken per day
```{r histogram, echo=TRUE, fig.height=6, fig.width=8}
#take the mean and median from the summed column
stepDayMean <- mean(stepDay$steps_by_day)
stepDayMedian <- median(stepDay$steps_by_day)
```
**The mean total number of steps taken per day**: 10766.1886792453  
**The median total number of steps taken per day**: 10765   

## What is the average daily activity pattern?  

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)  
```{r, echo=TRUE}
#average number of steps taken (mean) within each interval
stepInterval <- aggregate(list(stepMean=activity$steps), list(interval=activity$interval), mean, na.rm=TRUE)
#plot the line plot
plot(stepInterval, type="l", xlab="5-min Interval", ylab="Av Num of Steps", main = "Average Number of Steps Taken", col = "blue")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?  
```{r, echo=TRUE}
stepIntervalOrdered <- arrange(stepInterval, desc(stepMean))
stepIntervalOrdered[1,]
```
**The maximum number of steps (on average 206.1698) is on Interval 835.**  

## Imputing missing values  

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)  
```{r, echo=TRUE}
numNA <- sapply(activity, function(x) sum(is.na(x)))
numNA[[1]]
```
Source for the code is [here](http://stackoverflow.com/questions/8317231/elegant-way-to-report-missing-values-in-a-data-frame)  
**Total number of rows with NA:** 2304  

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.  
```{r, echo=TRUE}

```

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.  
```{r, echo=TRUE}

```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?  
```{r, echo=TRUE}

```

## Are there differences in activity patterns between weekdays and weekends?  
