---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data  

1. Unzip the file first, if the unzipped file is not already in the working directory.  
```{r echo = TRUE}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
```

2. Load the data (i.e. read.csv() )  
```{r echo = TRUE}
#read the table
activity <- read.csv("activity.csv", stringsAsFactors=FALSE, na.strings="NA")
```

3. Process/transform the data (if necessary) into a format suitable for your analysis  
```{r echo = TRUE}
#change the date column's class from string to Date
activity[,"date"] <- as.Date(activity[,"date"],format='%Y-%m-%d')
```

## What is mean total number of steps taken per day?  

1. Calculate the total number of steps taken per day  
```{r echo = TRUE}
library(dplyr)
activityDay <- group_by(activity, date)
stepDay <- summarize(activityDay, steps_by_day = sum(steps))
#For this part of the assignment, you can ignore the missing values in the dataset.
stepDay <- na.omit(stepDay)
```

Remove the NA's

## What is the average daily activity pattern?  



## Imputing missing values  



## Are there differences in activity patterns between weekdays and weekends?  
