---
title: "Reproducible Research: Peer Assessment 1"
author: "Nazmi Anik"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data  

1. Unzip the file first, if the unzipped file is not already in the working directory.  
```{r, echo = TRUE}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
```

2. Load the data ( i.e. read.csv() )  
```{r, echo = TRUE}
#read the table
activity <- read.csv("activity.csv", stringsAsFactors=FALSE, na.strings="NA")
```

3. Process/transform the data (if necessary) into a format suitable for your analysis  
```{r, echo = TRUE}
#change the date column's class from string to Date
activity[,"date"] <- as.Date(activity[,"date"],format='%Y-%m-%d')
#display the first few rows
head(activity)
```

## What is mean total number of steps taken per day?  

1. Calculate the total number of steps taken per day  
```{r, message=FALSE}
#import dplyr library
library(dplyr)
```
```{r, echo=TRUE}
#group the activity table by day
activityDay <- group_by(activity, date)
#summarize this grouped table with the sum of steps for each day
stepDay <- summarize(activityDay, steps_by_day = sum(steps))
#For this part of the assignment, you can ignore the missing values in the dataset.
stepDay <- na.omit(stepDay)
#display the first few rows
head(stepDay)
```

2. Make a histogram of the total number of steps taken each day  
```{r, echo=TRUE}
hist(stepDay$steps_by_day, main="Number of Steps Taken Each Day", xlab="Number of Steps per Day", ylab="Frequency", col = "red")
```

3. Calculate and report the mean and median of the total number of steps taken per day
```{r histogram, echo=TRUE, fig.height=6, fig.width=8}
#take the mean and median from the summed column
stepDayMean <- mean(stepDay$steps_by_day)
stepDayMedian <- median(stepDay$steps_by_day)
```
**The mean total number of steps taken per day**: 10766.1886792453  
**The median total number of steps taken per day**: 10765   

## What is the average daily activity pattern?  

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)  
```{r, echo=TRUE}
    
```
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?  
```{r, echo=TRUE}
    
```

## Imputing missing values  



## Are there differences in activity patterns between weekdays and weekends?  
